---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Using the activities.csv file found with our course, we'll load the data into a variable called 'data'.

The variables included in this dataset are:

1. steps: Number of steps taking in a 5-minute interval (missing values are coded as NA )
2. date: The date on which the measurement was taken in YYYY-MM-DD format
3. interval: Identifier for the 5-minute interval in which measurement was taken

```{r}
data = read.csv('activity.csv', header = T)
names(data)
```

Lets see what is in this dataset by using str

```{r}
str(data)

head(data)
```


## What is mean total number of steps taken per day?

Building a histogram of total number of steps taken each day. 

Using the aggregate function we'll ready the data for the histagram. Then we will produce the histogram.

1. Histogram
```{r}
step_date <- aggregate(steps ~ date, data = data, FUN = sum,)
barplot(step_date$steps, names.arg = step_date$date, xlab = "date", ylab = "steps")
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
